How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-29 tse-paper/content.tex
reformulate techniques argumentation, improve conclusion

\tableofcontents

@@

% TODO Caro: adapt as discussed
% TODO Caro: I think the axes labels are also unnecessary for this graph.
%Our other graphs in the lit survey do not have them, either.
\begin{figure}[tbhp]
		\centering
		\includegraphics[width=\columnwidth, trim={1.1cm 0.4cm
		1.5cm 0.5cm},
		clip]{img/lit-sur/log_producer_annotated.pdf}
		\caption{Frequency of supported log producers.}
		\label{fig:litsur:log_producer}
\end{figure}

@@

% Alt:
%\section{Analyzing Build Logs with Chunk Retrieval Techniques}
\section{Chunk Retrieval Techniques}
\label{sec:techniques}
%TODO: really? currently the discussion does not show this 'need'!
The outcome of our literature survey shows the need for a new
generation of techniques to analyze build logs.
In this section we argue how our research sets the first step to
address this need by introducing three techniques that retrieve
specific substrings (chunks) from build logs.
We present defining characteristics of these so called \emph{chunk
retrieval techniques} and present the techniques we developed in
detail.
At the end of this section we illustrate how the techniques
operate on a concrete
example and sketch out their implementation.

@@

\subsection{A New Generation of Build Log Analysis Techniques}
In summary, the results of the literature survey show the need for
build log
analysis techniques that are \emph{automated},
\emph{agnostic of the build environment}
and can be \emph{configured with little effort}.

% what we need is some derivation of the three techniques 1) regex 2)
%cts 3) kws, best repeated here in the discussion and then from the lit
%survey above, section 2.4.2
For this reason, we are focussing on techniques that run automatically
after being configured through examples or keywords.
\emph{Examples consist of logs and corresponding chunks.}
A user---a developer or a researcher who wants to analyze
build logs---provides these examples and keywords, which are easy to
adapt whenever the analysis needs to be maintained.

To analyze build logs,
we propose to use chunk retrieval techniques.
These retrieve a substring,
that describes a specific information, from a build log.
Generally, chunk retrieval techniques
can also be used to classify build logs, e.g.\ by checking
which substring was retrieved.
Therefore, chunk retrieval techniques
cover the great majority of the types of log analysis we saw within
our survey.

We implement three chunk retrieval techniques based on approaches
previously described by several articles:
PBE mimics regular expressions, while simplifying
their development.
We chose text similarity (CTS) as a common information retrieval
technique and implement the ad hoc approach of
searching for specific strings or keywords within KWS\@.

To evaluate and compare the three chunk retrieval techniques
we create a data set with logs from failed builds on Travis
CI---the most-used source of build logs within our survey.
For each log we marked the corresponding
chunks that describe error messages or
``the reason the build failed.''
This was the information
most targeted by the articles in our survey (see \Cref{sec:rq11}).
In addition to that, our data set encompasses a broad range of
build tools which enables us to measure the generic
applicability of the techniques.

% The articles within our study mainly tried to retrieve or classify
% errors or reasons for the build to fail from the logs.
% Therefore we chose the log chunk ``describing why the build failed''
% ~\cite{brandt2020logchunks} for our empirical comparison study
% of chunk retrieval techniques.
% The techniques we implement are in principle agnostic to the targeted
% log chunk, thus we believe our results can be generalized to other
% kinds of log chunks.

\subsection{Characteristics of Chunk Retrieval Techniques}
% TODO Caro: Wouldn't this place perhaps be better to contain
%the paragraph about why chunk log techniques are more general
% than other approaches?

\label{sec:crt-characteristics}
In this section, we formalize our use of chunk retrieval techniques.
By this word, we refer to techniques that automatically
retrieve pieces of information that appear literally in build
logs, i.e., techniques that do not aggregate, combine, or deduct
information.
We call such pieces of information in build logs
\emph{chunks}.
% and the techniques \emph{chunk retrieval techniques}.
The techniques we investigate here do not require a formal lexer and
parser to analyze the entire structure of build logs, but focus on
ad hoc extracting just one specific piece of information per
configuration.

@@

\subsubsection{Configuration}
A set of keywords configures the chunk retrieval with KWS\@.
To better
compare KWS with PBE and CTS, we also configure it through examples.
We associate a list of keywords with each example.
The keywords should appear in the targeted
chunk or close to it.
Then, we merge these lists and use the most commonly occuring
keywords for the search with KWS.

@@

\begin{subfigure}[tbp]{\columnwidth}
  \begin{lstlisting}[breaklines=true,frame=tlr]
=== RUN   TestSeparator
  \end{lstlisting}
  \vspace{-\baselineskip}
  \lstinputlisting[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]{listings/chunk2.txt}
  \vspace{-\baselineskip}
  % TODO Caro: Adapt frame= option to close frame properly
  \begin{lstlisting}[breaklines=true,frame=blr]
=== RUN   TestGenerateHTML
  \end{lstlisting}
  \caption{Example of a log chunk showing a linter error.}
  \label{lst:chunk-example-3}
\end{subfigure}

@@

In the following we will introduce the data set we created for this
study and describe the design of our comparison study.

\subsection{LogChunks Data Set}
\label{sec:logchunks}
To conduct the study in this paper, we created the
\emph{LogChunks} data set~\cite{brandt2020logchunks}.
It encompasses 797 build logs from Travis CI,
stemming from a broad range of 80 GitHub repositories and 29
programming languages.
For each file, we manually marked
the chunk that describes why the build
failed---the information targeted by most of the articles from our
literature survey.
To be able to configure the keyword search technique,
the data set associates each chunk with keywords that we
would use to search for the selected chunk within the log.
In
addition, we categorized the log chunks according to their format
within the log.
If the chunks were surrounded by the same markings
within the log, we assigned them the \emph{same structural category},
as described in \Cref{sec:crt-example}.

@@

	% actually that is now giving recall another (nicer) name
	% improved if we just talk about recall later?
	\item Successful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 1, \mathit{false\ otherwise}\vspace{0.1cm}
	\\$
	Partially Successful retrieval = $\mathit{true}\ \mathit{iff}\
	0 < \mathit{Recall} < 1, \mathit{false\ otherwise}\vspace{0.1cm}
	\\$
	Unsuccessful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 0, \mathit{false\ otherwise}\vspace{0.1cm}$

@@

The goal of this paper is to support practitioners and researchers in
their decision on how to analyze build logs.
To this end, we first
conducted a systematic mapping survey to map out the available build
log analysis techniques.
We saw that while build log analysis is widely used, it is often
a side part of an article and implementations are rarely reused.
This shows the need for a new generation of generic and automated
build log analysis techniques that need little effort to set up.
Based
on common approaches we saw during the literature survey,
we implemented three promising techniques to retrieve specific
substrings (chunks) from build logs:
PBE synthesizes regular expression programs based on examples of logs
and corresponding chunks, CTS leverages a common text similarity
measure to retrieve those lines which are most similar to example
chunks, and KWS finds the targeted parts of a build log by searching
for identifying keywords.
We evaluated the performance of these chunk retrieval techniques on the
\emph{LogChunks} data set and saw varying results.
CTS had the highest average F$_{1}$-score with 51\%, while the other
techniques showed higher average scores for precision (95\%, PBE)
and recall (70\%, KWS).
To guide researchers and developers on which technique to use for their
we summarized our results in a decision tree.
We found
that the structural representation of the targeted information in the
build logs is the main factor a user should consider when they
choose a suitable
technique.
Secondary factors are the desired confidence in recall and
precision of the produced output and whether precision or recall is
more important for the task at hand.
% TODO Caro: bring more across why it is so important that this field
% and the research there continues
Finally, we gave an
outline to guide future research efforts in the area of build log
analysis.
Important themes are a systematic investigation of industry approaches,
benchmarks of existing build log analysis approaches and further
refinement of the chunk retrieval techniques we proposed.

>>>>

@@

\begin{figure}[tbhp]
		\centering
		\includegraphics[width=\columnwidth, trim={1.5cm 0.4cm
		1.5cm 0.5cm},
		clip]{img/lit-sur/log_producer_annotated.pdf}
		\caption{Frequency of supported build tools.}
		\label{fig:litsur:log_producer}
\end{figure}

@@

% Alt:
%\section{Analyzing Build Logs with Chunk Retrieval Techniques}
\section{Chunk Retrieval Techniques}
\label{sec:techniques}

In summary, the outcome of the literature survey show the need for
build log
analysis techniques that are \emph{automated},
\emph{agnostic of the build environment}
and can be \emph{configured with little effort}.

To address this need,
we propose a set of so called \emph{chunk retrieval techniques},
which retrieve a specific substring (called chunk)
from a build log.
The techniques we propose run automatically
and can be easily configured through examples.
An example in this case consists of a build log together with the
chunk that should be retrieved.
A user---a developer or a researcher who wants to analyze
build logs---provides these examples, which are easy to
adapt whenever the analysis technique needs
to support new cases.

We designed and implemented three chunk retrieval
techniques based on approaches
previously described by several articles:
PBE mimics regular expressions, while simplifying
their development.
We chose text similarity (CTS) as a common information retrieval
technique and implement the ad hoc approach of
searching for specific strings or keywords within KWS\@.

In this section, we define our notion of chunk
retrieval techniques and present each of the techniques
we developed in detail.
Following that, we illustrate how the techniques
operate on a concrete
example and sketch out their implementation.

@@

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:crt-characteristics}
In this section, we formalize our use of chunk retrieval techniques.
By this word, we refer to techniques that automatically
retrieve pieces of information that appear literally in build
logs, i.e., techniques that do not aggregate, combine, or deduct
information.
We call such pieces of information in build logs
\emph{chunks}.

Generally, chunk retrieval techniques
can also be used to classify build logs, e.g.\ by checking
which substring was retrieved.
Therefore, chunk retrieval techniques
cover the great majority of the types of log analysis we saw within
our survey.

The techniques we investigate here do not require a formal lexer and
parser to analyze the entire structure of build logs, but focus on
ad hoc extracting just one specific piece of information per
configuration.

@@

\subsubsection{Configuration}
A set of keywords configures the chunk retrieval with KWS\@.
To better
compare KWS with PBE and CTS, we also configure it through examples.
We associate a list of keywords with each example.
The keywords should appear in the targeted
chunk or close to it.
Then, we merge these lists and use the most commonly occurring
keywords for the search with KWS.

@@

\begin{subfigure}[tbp]{\columnwidth}
  \begin{lstlisting}[breaklines=true,frame=tlr]
=== RUN   TestSeparator
  \end{lstlisting}
  \vspace{-\baselineskip}
  \lstinputlisting[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]{listings/chunk2.txt}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]
=== RUN   TestGenerateHTML
  \end{lstlisting}
  \caption{Example of a log chunk showing a linter error.}
  \label{lst:chunk-example-3}
\end{subfigure}

@@

In the following, we introduce the data set we created for this
study and describe the design of our comparison study.

\subsection{LogChunks Data Set}
\label{sec:logchunks}
To conduct the study in this paper, we created the
\emph{LogChunks} data set~\cite{brandt2020logchunks}.
From Travis CI---the most-used source of build logs within
according to our literature survey---we collected
797 build logs that
stem from a broad range of 80 GitHub repositories and 29
programming languages.
This wide variety of projects, and the large number of different build
tools they employ, enables us to measure the generic
applicability of the chunk retrieval techniques.

For each build log, we manually marked
the chunk that describes why the build
failed---the information targeted by most of the articles in our
literature survey.
To be able to configure the keyword search technique,
the data set associates each chunk with keywords that we
would use to search for the selected chunk within the log.
In
addition, we categorized the log chunks according to their format
within the log.
If the chunks were surrounded by the same markings
within the log, we assigned them the \emph{same structural category},
as described in \Cref{sec:crt-example}.

@@

	% actually that is now giving recall another (nicer) name
	% improved if we just talk about recall later?
	\item Successful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 1, \vspace{0.1cm}\\$
	Partially successful retrieval = $\mathit{true}\ \mathit{iff}\
	0<\mathit{Recall}<1, \vspace{0.1cm}\\$
	Unsuccessful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 0, \vspace{0.1cm}$

@@

The goal of this paper is to support practitioners and researchers in
their decision on how to analyze build logs.
To this end, we
conducted a systematic mapping survey to map out the available build
log analysis techniques.
We saw that, while build log analysis is widely used, it is often
a side part of an article and implementations are rarely reused.
This shows the need for a new generation of generic and automated
build log analysis techniques that need little effort to set up.

Based on common approaches we observed in literature,
we implemented three promising chunk retrieval techniques:
PBE synthesizes regular expression programs based on examples of logs
and corresponding chunks, CTS leverages a common text similarity
measure to retrieve those lines which are most similar to example
chunks, and KWS finds the targeted chunks in a build log by searching
for identifying keywords.

We evaluated the performance of these chunk retrieval techniques on the
\emph{LogChunks} data set and saw varying results.
CTS had the highest average F$_{1}$-score with 51\%, while the other
techniques showed higher average scores for precision (95\%, PBE)
and recall (70\%, KWS).

To guide researchers and developers on which technique suits their
use case,
we summarized our results in a decision tree.
We found
that the structural representation of the targeted information in the
build logs is the main factor a user should consider when they
choose a
technique.
Secondary factors are the desired confidence in recall and
precision of the produced output and whether precision or recall is
more important for the task at hand.

Finally, we gave an
outline to guide future research efforts in the area of build log
analysis.
Important themes are a systematic investigation of industry approaches,
benchmarks of existing build log analysis approaches and further
refinement of the chunk retrieval techniques we proposed.
